---
title: 'Case Study 2: How Can a Wellness Technology Company Play It Smart?'
author: "Thuan Ha"
date: "2024-01-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction:
This is a case study from the Google Data Analytics Course on Coursera, where I
will be using what I have learn throughout the course to perform a real-world 
tasks of a junior data analyst. 

# Scenario:
You are a junior data analyst working on the marketing analyst team at 
Bellabeat, a high-tech manufacturer of health-focused products for women. 
Bellabeat is a successful small company, but they have the potential to become 
a larger player in the global smart device market. Urška Sršen, co-founder and 
Chief Creative Officer of Bellabeat, believes that analyzing smart device 
fitness data could help unlock new growth opportunities for the company. You 
have been asked to focus on one of Bellabeat’s products and analyze smart 
device data to gain insight into how consumers are using their smart devices. 
The insights you discover will then help guide marketing strategy for the 
company.

# Dataset:
The dataset provide is FitBit Fitness Tracker Data from Kaggle. This Kaggle data 
set contains personal fitness tracker from thirty fitbit users. Thirty eligible 
Fitbit users consented to the submission of personal tracker data, including 
minute-level output for physical activity, heart rate, and sleep monitoring. It 
includes information about daily activity, steps, and heart rate that can be 
used to explore users’ habits.

# Bussiness Tasks:
1. What are some trends in smart device usage?/
2. How could these trends apply to Bellabeat customers?/
3. How could these trends help influence Bellabeat marketing strategy?/

## Packages use for this study:

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(skimr)
library(dbplyr)
library(stringr)
```

## Importing the data:

```{r}
daily_activity <- read.csv("Bellabeat Data/dailyActivity_merged.csv")
daily_sleep <- read.csv("Bellabeat Data/sleepDay_merged.csv")
weight_log_info <- read.csv("Bellabeat Data/weightLogInfo_merged.csv")
hourly_intensities <- read.csv("Bellabeat Data/hourlyIntensities_merged.csv")
```

### Checking the data:
Using the head function let me have a quick overview of the data to make sure 
each columns have the correct format. It look like the date columns for the 
four dataset is in character format instead of date format, and IsManualReport 
for the weight_log_info dataset could be change to a Boolean format.

```{r}
head(daily_activity)
head(daily_sleep)
head(weight_log_info)
head(hourly_intensities)
```

## Cleaning columns:
I'll be using the clean_names() function to have a consistent columns name for
the dataset I'll be using and to remove space, letter cases, etc. to a 
consistent format. Also, I'll be changing the date format and adding new columns
that could help with my analysis.

```{r}
daily_activity <- clean_names(daily_activity)
daily_sleep <- clean_names(daily_sleep)
weight_log_info <- clean_names(weight_log_info)
hourly_intensities <- clean_names(hourly_intensities)
```

```{r}
n_distinct(weight_log_info$id)
```

After looking over the data I decide to check the weight_log_info dataset since 
it contain many null value in one of the columns. It seem there are only 8 
participant willing to give their weight info which is hard to have a analysis 
for the weight dataset.


```{r}
#as.Date function let me convert chr column into date
daily_activity$activity_date <- as.Date(daily_activity$activity_date, "%m/%d/%y")
```

```{r}
daily_sleep$sleep_day <- as.Date(daily_sleep$sleep_day, "%m/%d/%y")
```

```{r}
hourly_intensities$activity_hour <- parse_date_time(hourly_intensities$activity_hour,
                                               "%m/%d/%y %H:%M:%S %p")
```


```{r}
#for this dataset I use parse_date_time instead since it can handle the AM/PM in the date
weight_log_info$date <- parse_date_time(weight_log_info$date, "%m/%d/%y %H:%M:%S %p")
```

```{r}
#the is_manual_report contain True or False so I change it into a Boolean
weight_log_info$is_manual_report <- as.logical(weight_log_info$is_manual_report)
```



### Adding new columns:
Adding a total active hours and days of week columns to help further analyze 
trend such as which day are participants most active.

```{r}
daily_activity <- daily_activity %>%
  mutate(total_active_hours = round((very_active_minutes + fairly_active_minutes + lightly_active_minutes)/60),
         days_of_week = wday(activity_date, label = T))
```

```{r}
hourly_intensities$Time <- format(as.POSIXct(hourly_intensities$activity_hour,format="%Y:%m:%d %H:%M:%S"),"%H:%M:%S")

hourly_intensities$Date <- format(as.POSIXct(hourly_intensities$activity_hour,format="%Y:%m:%d %H:%M:%S"),"%Y:%m:%d")
```


```{r}
daily_activity_cleaned <- daily_activity[!(daily_activity$total_active_hours <= 0.00),]
```


### Dataset summary:

```{r}
# daily activiy summary
daily_activity %>%
  select(total_steps, 
         total_distance, 
         total_active_hours) %>%
  summary()

# summary of active by minutes
daily_activity %>%
  select(very_active_minutes, 
         fairly_active_minutes, 
         lightly_active_minutes) %>%
  summary()

# calories and minutes spend sitting
daily_activity %>%
  select(sedentary_minutes, 
         calories) %>%
  summary()

# time spend sleeping
daily_sleep %>%
  select(total_minutes_asleep, 
         total_time_in_bed) %>%
  summary()
```

### Info. from the summary:

* The most stand out observation would be the average sedentary time or time 
spend sitting is 991 minutes or around 16-17 hours, which is consider a high
risk according to this article.[linked phrase](https://www.medicalnewstoday.com/articles/sitting-down-all-day#how-long-is-too-long)/

* The average total steps is 7,638 which is under the recommended average of
10,000 steps by the CDC.[linked phrase](https://www.medicalnewstoday.com/articles/how-many-steps-should-you-take-a-day#:~:text=As%20a%20result%2C%20the%20CDC,to%20about%201.5%E2%80%932%20miles.)/

* While the average active hours is around the estimated of 4 hours, majority 
came from lightly active. It is recommended that adults aged between 18 - 64 
years should do at least 150–300 minutes of moderate-intensity physical 
activity.[linked phrase](https://www.who.int/news-room/fact-sheets/detail/physical-activity)/

## Visualization:

```{r}
daily_activity %>%
  group_by(days_of_week) %>%
  summarise(average_total_active_hours = mean(total_active_hours)) %>%
  ggplot(aes(x = days_of_week, y = average_total_active_hours)) +
  geom_col(fill = "blue") +
  labs(title = "Average active hours per day")
```

```{r}
daily_activity %>%
  group_by(days_of_week) %>%
  ggplot(aes(x = days_of_week, y = total_steps)) +
  geom_col(fill = "blue") +
  labs(title = "Total steps per day")
```

```{r}
daily_activity %>%
  group_by(days_of_week) %>%
  ggplot(aes(x = days_of_week, y = total_active_hours)) +
  geom_col(fill = "blue") +
  labs(title = "Total active hours per day")
```

```{r}
daily_activity %>%
  group_by(days_of_week) %>%
  ggplot(aes(x = days_of_week, y = calories)) +
  geom_col(fill = "blue") +
  labs(title = "Total calories burned per day")
```



```{r}
avg_int_by_hour <- hourly_intensities %>%
  group_by(Time) %>%
  drop_na() %>%
  summarise(average_total_intensity = mean(total_intensity))

ggplot(avg_int_by_hour, aes(x = Time, y = average_total_intensity)) +
  geom_histogram(stat = "identity", fill = "dark blue") +
  theme(axis.text.x = element_text(angle = 270)) +
  labs(title = "Time when there are most intensisies")
```


```{r}
aggregate(daily_activity$total_active_hours, list(daily_activity$days_of_week), FUN = sum)
```


```{r}
daily_activity %>%
  group_by(days_of_week) %>%
  summarise(average_sedentary_minutes = mean(sedentary_minutes)) %>%
  ggplot(aes(x = days_of_week, y = average_sedentary_minutes)) +
  geom_col(fill = "blue") +
  labs(title = "Average time spent sitting per days")
```
### Correlations:

```{r}
cor(daily_activity$total_steps, daily_activity$calories)
```

```{r}
cor(daily_activity$total_active_hours, daily_activity$calories)
```

```{r}
cor(daily_activity$very_active_minutes, daily_activity$calories)
```


```{r}
cor(daily_activity$sedentary_minutes, daily_activity$calories)
```


```{r}
ggplot(daily_activity, aes(x = total_steps, y = calories)) + geom_smooth() + geom_point(color = "maroon")
```

```{r}
ggplot(daily_activity, aes(x = total_active_hours, y = calories)) + geom_jitter(color = "maroon") + geom_smooth() + labs(title = "Correlations between calories burned and active hours")
```

```{r}
ggplot(daily_activity, aes(x = very_active_minutes, y = calories)) + geom_jitter(color = "maroon") + geom_smooth() + labs(title = "Very active minutes vs. calories burn")
```

```{r}
ggplot(daily_activity, aes(x = lightly_active_minutes, y = calories)) + geom_jitter(color = "maroon") + geom_smooth() + labs(title = "Light active minutes vs. calories burn ")
```


```{r}
ggplot(daily_activity, aes(x = sedentary_minutes, y = calories)) + geom_jitter(color = "maroon") + geom_smooth() + labs(title = "Correlations between calories burned and sedentary")
```

## Summary
After analyzing the plot, it seem that there are a drop off in active hour after
sunday, while the average active hours are around 3-4 hours a day from the first
plot we know that it mostly consist of lightly active from a quick summary that
we did before it./

Also, there are a strong correlation between total steps and calories burn from 
the correlation test and the plot. There are correlation between active hours 
and calories burn, however there are only moderate since the correlation 
coefficient is only 0.46. The relationship between active and calories burn does
get better as we look at the correlation test and plot between very active and
lightly active./

### Recommendation and ideas
1. Bellabeat could have recommend workout routine through the app that let user 
know what workout they do for each day. From a personal experience, when I first
start workout it very demotivating since I have no idea what to do, which in
turn cause me to be lazy and not go to the gym, but after having a workout 
routine it give me no excuse since I have a set workout ready for me./

2. Having a notification of calories burn after a workout also motivate an
individual since it show a goal that they have achieve./

3. The average sedentary is very concerning, what Bellabeat could do is have
notification through there app that let user know when the sedentary time is 
high and recommend to do a light walk or stand up to stretch./


